Compositional equivalences based on Open pNets

### 2020-02-12 2016-Weak-FH-Bisim/WeakBisim.tex
tr2, lemma2, double qed

	The rule for a pLTS  checks that the guard 
	is verified and transforms assignments into post-conditions:

	\begin{description}
		\item[{\TrUn:}]
		$\inferrule
		{ s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
					J}\rangle} s'\in \to  }
		{ \mylangle  S,s_0, \to \myrangle
			\models
			\openrule
			{\emptyset ,
			e_b,\left\{x_j\gets e_j\right\}^{j\in J}}
			{\ostate{s} \OTarrow{\alpha} \ostate{s'}}
		}
		$
	\end{description}
%	Note that this note is greatly simplified by the fact that variables are local to 
%	thread; introducing global state variables or accepting loops to the same 
%	state would 
%	require to reason 
%	on the scope of 
%	each variables, and to introduce additional variables to handle the several occurence 
%	of the same pLTS variable in the predicates. Indeed the constraints on pLTS 
%	transitions 
%	ensure that the same variable never appears both on the left and on the right of the 
%	equations of a predicate.

	The second rule deals with pNet nodes: for each possible
	synchronisation vector (of index $k$) applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved , one possible
	{\em action} for each Hole involved, and the predicate relating these
	with the resulting action of the vector. The sub-pNets involved are split between two 
	sets, $I_2$ for subnets that are pLTSs, and $I_1$ for the others, $J$ is the set of 
	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{({\alpha'})_m^{m 
	\in M}}{\alpha'}{e_b}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
	I_2$}.

	\begin{description}
		\item[{\TrDeux:}]
	\end{description}

	\noindent
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
    J' = \biguplus_{m\in I_1}\!\! J_m \uplus J	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j^{j\in J},\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		\beta_j^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}    
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}
\end{definition}
        	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.

@@

We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:
\begin{lemma}[Open transition decomposition]\label{lem-decompose} Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

Lemma \ref{lem-compose} is combining an open transition of $P$ with
an open transition of $Q$, and building a corresponding transition of
$P[Q]_{j_0}$, assembling their predicates.
\paragraph{Lemma~\ref{lem-compose}} {\bf Open transition composition}\\
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]

@@

@@

   \begin{proof}
Consider rule \TrDeux\ in 
       Definition~\ref{def:operationalSemantics}, applied to the pNet $P[Q]_{j_0}$.

\noindent\TrDeux:\\
	\noindent
\begin{small}	
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}

@@

@@

       Consider each premise of the open transition (constructed by \TrDeux\ in 
Definition~\ref{def:operationalSemantics}). Let $P=\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle$
	\begin{description}
		\item[{\TrDeux:}]
	\end{description}

	\noindent

@@

@@

	\noindent
	\begin{small}
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I\cup\{j_0\}}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1\cup\{j_0\}. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J'_{PQ}=  J\setminus\{j_0\}\uplus J_Q	\\
    	\Pred\,'' =\bigwedge_{m\in I _1\uplus I_2}\Pred_m \land \Pred\,'
\land \Pred(SV_k,v_i^{i\in I_1\uplus I_2}\uplus(j_0\mapsto v_Q),\beta_j^{j\in 
J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1\cup\{j_0\}}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in  J'_{PQ}}, \Pred\,'',  	\Post\uplus \Post\,'}
    		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L\uplus L_Q}}}
    	}
    }
\end{mathpar}  
\end{small}
 \qed
        \end{proof}

@@

 Note that, $\beta_{j_0}$ can be transformed into  $\beta_{j_0 x}$ because of the 
 implication hypothesis.
 %\TODO{There was a confusion between the $\Rightarrow$ between the
 %  steps, and the $\implies$ inside the formulas...}
% 
% Explanations: first, we reversed the substitution; then we
% restricted the 
% substitution as $\beta_{j_0}$ only appears in $\Pred\,'$; then we
% applied the property of FH-bisimulation for $P$ and rearranged the
% terms using the fact that $v_Q=v_x$.
 The obtained formula reaches the goal except for 
 two points:\\[-4.3ex] 
 \begin{itemize}
 	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J\cap\Holes(Q)$  but  
 	adding prerequisite on more variables 
 	does not   	change the validity of the formula (those variables are not used).
 	\item Concerning the last term, we need 
 	$\Pred_{s x}\subst{\Post_{OT}\uplus(\Post\,'\uplus \Post_{OT_x})}$, i.e.
 	$\Pred_{s, x}\subst{(\Post\,'\uplus 	\Post\,'')\uplus(\Post\,'\uplus \Post_{OT_x})}$. We 
 	can conclude by observing that	$\Pred_{s,x}$ does not use any variable of $P$ 
 	and thus the substitution $\subst{Post\,'}$ has no effect on it.
 \end{itemize}	
Finally: \\
\begin{small} $\Pred_{s,t} \land \Pred_{OT} \implies\\
{\displaystyle\bigvee_{x\in X}
 		\!\left( \forall j\in J. \beta_j=\beta_{jx} \! \land\! 
 		\big(\Pred\,'\land\! \Pred_{OT_x}
 		 \land \alpha_Q\!=\!\beta_{j_0 x}\right)\! \land\! \alpha_Q\!=\!\alpha_x\land\!  \Pred_{s, x}\subst{\Post\,''\!\uplus\!\Post_{OT_x}}\big)
 		}$
 		\end{small}\\
This proves the  condition of the FH-simulation, the other direction is 
 similar.\qed

@@

Also, we have the following assertions:\\
$
{\displaystyle \alpha''=\alpha\subst{\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4j}}\!}\\
s\!=\!s_{40} \wedge s_{4(n_1+n_2+1)}\!=s_2 \wedge s_2' = s_{50} \wedge s_{5(m_1+m_2+1)}\!=\!s'\\
~~\\
\displaystyle{\set{\gamma}=\mybigdotcup_{i=0}^{n_1+n_2} \vis{\set{\beta_{4i}}\subst{\bigotimes_{j=i-1}^{0} \Post_{4 j} } }  \dotcup  \vis{\set{\beta'}\subst{\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4j}}}\dotcup }\\~\!\qquad\mybigdotcup_{i=0}^{m_1+m_2}\!\!\vis{\set{\beta_{5i}}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{5i}\shortotimes\Post\,'\shortotimes\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4i}} }
\\
\displaystyle{\Pred=
\bigwedge_{i=0}^{n_1+n_2}\!\!\Pred_{4i}\subst{\bigotimes_{j=i-1}^{0}\Post_{4j}}\land\Pred\,' \subst{\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\!\Post_{4j}}\land}\\ 
~\qquad\quad\bigwedge_{i=0}^{m_1+m_2}\Pred_{5i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{5j}\shortotimes\Post\,'\shortotimes\!\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\!\Post_{4j}}\\
\Post=\bigotimes_{j=m_1+m_2}^{0}\!\!\!\Post_{5j}\shortotimes\Post\,'\shortotimes\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\Post_{4j}
$\\
This concludes the inductive step, showing  that the decomposition expressed by the $\Leftarrow$ direction of the lemma is always possible with the right side conditions. \qed
\end{itemize}
\end{proof}

@@

\qed

>>>>

	The rule for a pLTS  checks that the guard 
	is verified and transforms assignments into post-conditions:		
\begin{mathpar}\inferrule
		{ s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
					J}\rangle} s'\in \to  }
		{ \mylangle  S,s_0, \to \myrangle
			\models
			\openrule
			{\emptyset ,
			e_b,\left\{x_j\gets e_j\right\}^{j\in J}}
			{\ostate{s} \OTarrow{\alpha} \ostate{s'}}
		}\quad {\TrUn}
\end{mathpar}
%	Note that this note is greatly simplified by the fact that variables are local to 
%	thread; introducing global state variables or accepting loops to the same 
%	state would 
%	require to reason 
%	on the scope of 
%	each variables, and to introduce additional variables to handle the several occurence 
%	of the same pLTS variable in the predicates. Indeed the constraints on pLTS 
%	transitions 
%	ensure that the same variable never appears both on the left and on the right of the 
%	equations of a predicate.

	The second rule deals with pNet nodes: for each possible
	synchronisation vector (of index $k$) applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved , one possible
	{\em action} for each Hole involved, and the predicate relating these
	with the resulting action of the vector. The sub-pNets involved are split between two 
	sets, $I_2$ for subnets that are pLTSs, and $I_1$ for the others, $J$ is the set of 
	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{({\alpha'})_m^{m 
	\in M}}{\alpha'}{e_b}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
	I_2$}.                                                                    
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
    J' = \biguplus_{m\in I_1}\!\! J_m \uplus J	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j^{j\in J},\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		\beta_j^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }\quad {\TrDeux}
\end{mathpar}    
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}
\end{definition}
        	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.

@@

We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:
\begin{lemma}[Open transition decomposition\label{lem-decompose}] Consider two pNets $P$ and $Q$ that are not pLTSs\footnote{A similar lemma can be proven for a pLTS $Q$}.
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\beta_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.
		 Then there exist $\alpha_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\beta_j^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\openrule
			{
				\beta_j^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred \iff \Pred\,'
		\land \Pred\,''\land \alpha_Q=\beta_{j_0}$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

Lemma \ref{lem-compose} is combining an open transition of $P$ with
an open transition of $Q$, and building a corresponding transition of
$P[Q]_{j_0}$, assembling their predicates.

@@

\TODO{LUDO: paragraphe d intro et de decription scientifique}

@@

   \begin{proof}
Consider rule \TrDeux\ in 
       Definition~\ref{def:operationalSemantics}, applied to the pNet $P[Q]_{j_0}$. 	
	\noindent
\begin{small}	
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }\quad\TrDeux
\end{mathpar}

@@

\paragraph{\bf Lemma~\ref{lem-compose}} {\it Open transition composition}\\
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]

@@

       Consider each premise of the open transition (constructed by \TrDeux\ in 
Definition~\ref{def:operationalSemantics}). Let $P=\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle$

@@

    }\quad \TrDeux

@@

	\noindent
	\begin{small}
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I\cup\{j_0\}}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1\cup\{j_0\}. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J'_{PQ}=  J\setminus\{j_0\}\uplus J_Q	\\
    	\Pred\,'' =\bigwedge_{m\in I _1\uplus I_2}\Pred_m \land \Pred\,'
\land \Pred(SV_k,v_i^{i\in I_1\uplus I_2}\uplus(j_0\mapsto v_Q),\beta_j^{j\in 
J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1\cup\{j_0\}}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in  J'_{PQ}}, \Pred\,'',  	\Post\uplus \Post\,'}
    		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L\uplus L_Q}}}
    	}
    }
\end{mathpar}  
\end{small}
        \end{proof}

@@

 Note that, $\beta_{j_0}$ can be transformed into  $\beta_{j_0 x}$ because of the 
 implication hypothesis.
 %\TODO{There was a confusion between the $\Rightarrow$ between the
 %  steps, and the $\implies$ inside the formulas...}
% 
% Explanations: first, we reversed the substitution; then we
% restricted the 
% substitution as $\beta_{j_0}$ only appears in $\Pred\,'$; then we
% applied the property of FH-bisimulation for $P$ and rearranged the
% terms using the fact that $v_Q=v_x$.
 The obtained formula reaches the goal except for 
 two points:
 \begin{itemize}
 	\item We need $\forall j\!\in\! J$ instead of $\forall j\!\in\! J\cap\Holes(Q)$  but  
 	adding prerequisite on more variables 
 	does not   	change the validity of the formula (those variables are not used).
 	\item Concerning the last term, we need 
 	$\Pred_{s x}\subst{\Post_{OT}\uplus(\Post\,'\uplus \Post_{OT_x})}$, i.e.
 	$\Pred_{s, x}\subst{(\Post\,'\uplus 	\Post\,'')\uplus(\Post\,'\uplus \Post_{OT_x})}$. We 
 	can conclude by observing that	$\Pred_{s,x}$ does not use any variable of $P$ 
 	and thus the substitution $\subst{Post\,'}$ has no effect on it.
 \end{itemize}	
Finally: \\
\begin{small} $\Pred_{s,t} \land \Pred_{OT} \implies\\
{\displaystyle\bigvee_{x\in X}
 		\!\left( \forall j\in J. \beta_j=\beta_{jx} \! \land\! 
 		\big(\Pred\,'\land\! \Pred_{OT_x}
 		 \land \alpha_Q\!=\!\beta_{j_0 x}\right)\! \land\! \alpha_Q\!=\!\alpha_x\land\!  \Pred_{s, x}\subst{\Post\,''\!\uplus\!\Post_{OT_x}}\big)
 		}$
 		\end{small}\\
This proves the  condition of the FH-simulation, the other direction is 
 similar.\qed

@@

Also, we have the following assertions:\\
$
{\displaystyle \alpha''=\alpha\subst{\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4j}}\!}\\
s\!=\!s_{40} \wedge s_{4(n_1+n_2+1)}\!=s_2 \wedge s_2' = s_{50} \wedge s_{5(m_1+m_2+1)}\!=\!s'\\
~~\\
\displaystyle{\set{\gamma}=\mybigdotcup_{i=0}^{n_1+n_2} \vis{\set{\beta_{4i}}\subst{\bigotimes_{j=i-1}^{0} \Post_{4 j} } }  \dotcup  \vis{\set{\beta'}\subst{\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4j}}}\dotcup }\\~\!\qquad\mybigdotcup_{i=0}^{m_1+m_2}\!\!\vis{\set{\beta_{5i}}\subst{\bigotimes_{j=i-1}^{0}\Post\,'_{5i}\shortotimes\Post\,'\shortotimes\!\!\bigotimes_{j=n_1+n_2}^{0}\!\!\!\Post_{4i}} }
\\
\displaystyle{\Pred=
\bigwedge_{i=0}^{n_1+n_2}\!\!\Pred_{4i}\subst{\bigotimes_{j=i-1}^{0}\Post_{4j}}\land\Pred\,' \subst{\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\!\Post_{4j}}\land}\\ 
~\qquad\quad\bigwedge_{i=0}^{m_1+m_2}\Pred_{5i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{5j}\shortotimes\Post\,'\shortotimes\!\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\!\Post_{4j}}\\
\Post=\bigotimes_{j=m_1+m_2}^{0}\!\!\!\Post_{5j}\shortotimes\Post\,'\shortotimes\!\!\bigotimes_{j={n_1+n_2}}^{0}\!\!\Post_{4j}
$\\
This concludes the inductive step, showing  that the decomposition expressed by the $\Leftarrow$ direction of the lemma is always possible with the right side conditions. 
\end{itemize}
\end{proof}

@@

%\qed
